---
title: "The Leslie Matrix - revisited"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A population of fish we're interested in has four life stages: eggs (E), fry (F), juvenile (J), breeding adult (A). You are told that annually:

- Each breeding adult will produce, on average, 600 eggs
- 5% of eggs will survive to become fry
- 20% of fry will survive to become juveniles
- 16% of juveniles will survive to become adults
- Adult survival rate year to year is 90%

Recall the Leslie matrix from EDS 212. Matrix multiplication (using the dot product `%*%` of the matrix with the populations at each life stage) yields the population structure at the following year. 

### Create the Leslie matrix:

```{r}
# Make the Leslie Matrix:
fish_leslie <- matrix(c(0, 0, 0, 600, 0.05, 0, 0, 0, 0, 0.20, 0, 0, 0, 0, 0.16, 0.9), nrow = 4, ncol = 4, byrow = TRUE)

# Check it out: 
fish_leslie
```

The initial population structure is given by: 
```{r}
# Initial population structure (Roe, Fry, Juvenile, Adult): 
fish_year0 <- c(5000, 8000, 600, 400)
```


Add line-by-line comments wheres prompted in the code below, explaining what each is doing. Make sure that you are talking with your neighbors about what is really being created at each step. 

```{r}
# Model projections

# Creating a sequence from 0 to 8 w increments of 1 and assigned it years
time_yr <- seq(from = 0, to = 8, by = 1)

# Create a matrix using fish years intial and rows populating by year- 9 rows , 4 cols
proj_year <- matrix(ncol = length(fish_year0), nrow = length(time_yr))

# Populate rows first by fish years(initial), "this is saying i want row 1 of proj year matrix to be populated by the initial condition of fish_year0" -allison
proj_year[1, ] <- fish_year0

#For loop demonstrating to pull the time_yr vector and populate it with projected years with rows first and assigning that with the corresponding col that will multiply but the projected year (-1- so populating first growth)
for (i in 2:length(time_yr)) { #first population is year so she sets it to 2
  proj_year[i,] <- fish_leslie %*% proj_year[i-1,] # dot product of that matrix of coefficients (that tell me expected change) and the row containing the population structure data from the previous cycle
}

# The rest is wrangling & visualization (run to check it out): #reassigning of porj year to :
colnames(proj_year) <- c("eggs", "fry", "juvenile", "adult")

#Pivot longer across all cols except "-time_yr" and take all col names and put those into single cols and take all values of thos cols and move it to a new one called stage population
proj_df <- data.frame(time_yr, proj_year) %>% 
  pivot_longer(cols = -time_yr, names_to = "lifestage", values_to = "stage_population")

ggplot(data = proj_df, aes(x = time_yr, y = stage_population)) +
  geom_line(aes(color = lifestage)) +
  scale_y_log10()

```

Test different combinations of survival / reproduction and starting population structures to see what happens! 

```{r}
demo <- data.frame(year = c(1995, 2000, 2005),
                   cats = c(2, 4, 4),
                   dogs = c(10, 14, 12))

ggplot(data = demo) + 
  geom_line(aes(x = year, y = cats)) +
  geom_line(aes(x = year, y = dogs)) 


demo_longer <- demo |>
  pivot_longer(cols = c(cats, dogs), names_to = "species", values_to = "count")
```

